Fréchet Distance

### 2022-11-22 docs/paper.tex
add Fréchet distance to intro section

%%%
%%% Colors
%%%

\RequirePackage[table,dvipsnames]{xcolor}

\definecolor{TUGreen}{rgb}{0.517,0.721,0.094}

@@

\section{Introduction}
% motivation
At first it seems that measuring the similairty between two curves
is a very theoretical problem
but it has many, not that obvious, pratical applications.
It is used in Geo-Information-Systems (GIS) to 
quantify the similarity or quality of two maps, % cite
find patterns in vehicle traffic or animal tracking data 
by trajectory analysis, % cite
gesture recognition on touch devices like smartphones, % cite
and shape matching in the field of computer vision.

@@

% two (polygonal) curves
% TODO: replace continous funcitons with other name
In the following we'll focus on polygonal curves (polygonal chains) 
instead of continuous functions ofen used in pure mathematics 
because most data we have for these problems is 
captured by some none continuous mechanisms, e.g., GPS positions from a tracking device.
Therefore we consider a polygonal curve $P$ to be definded as a 
sequence of points $p_1, \dots, p_n \in R^d$ where $d$ is the dimensionality 
of the space, e.g., $n=2$ or $n=3$, and linear interpolation between this points. 
However, many concepts also apply to continuous functions.

% Hausdorff distance
The commonly known \textit{Hausdorff distance} can be understood 
as the largest distance between any two points on the curves. 
For each curve, a point with the maximum shortest distance to a point on the other curve is searched for, see \autoref{fig_hausdorff_distance_example}.
The Hausdorff distance is mathematically formulated as stated in \autoref{eq_hausdorff_distance} \cite{alt_computing_1995}.

\begin{align}
  \label{eq_hausdorff_distance}
  \delta_{hd}(P,Q) = \max \left( \sup_{p \in P} \inf_{q \in Q} \text{dist}(p,q), \sup_{q \in Q} \inf_{p \in P} \text{dist}(p,q) \right)
\end{align}

However, this metric does not take the course of the curves into account. 
Two curves may have a small Hausdorff-distance but seem very dissimilar 
when looking at them, see \autoref{fig_hausdorff_distance_bad_example}.

@@

% Fréchet distance motivation

@@

% Scope

% Fréchet distance Variants 
\section{Variants}
Through the years many slightly differnt variants of the continuous Fréchet distance have evolved. 
The continuous Fréchet distance takes into account every single point on the curves, 
with interpolation in the case of polygonal curves, 
without allowing to walk backwards 
and without any obstacles in between the two curves.

>>>>

%%%
%%% Colors
%%%
\RequirePackage[table,dvipsnames]{xcolor}
\definecolor{TUGreen}{rgb}{0.517,0.721,0.094}

%%%
%%% Macros
%%%
\newtheorem{mydef}{Definition}

@@

\section{Introduction}
% motivation
At first it seems that measuring the similairty between two curves
is a very theoretical problem
but it has many, not that obvious, pratical applications.
It is used in Geo-Information-Systems (GIS) to 
quantify the similarity or quality of two maps, % cite
find patterns in vehicle traffic or animal tracking data 
by trajectory analysis, % cite
gesture recognition on touch devices like smartphones \cite{hu_research_2022}, % cite
and shape matching in the field of computer vision.

@@

% two (polygonal) curves
% TODO: replace continous funcitons with other name
\begin{mydef}
  \label{def_curve}
  We define a curve as a continuous function $f : [a,b] \mapsto V$ 
  with $a,b \in \mathbb{R}, a \leq b$ and $V$ being the space of the curve,
  e.g., $\mathbb{R}^2$ or $\mathbb{R}^3$.
\end{mydef}

\begin{mydef}
  \label{def_polygonal_curve}
  We define a polygonal curve (polygonal chain) 
  as a linearly interpolated curve 
  on $n \in \mathbb{N}$ points in $V$ such that
  $P : [0,n] \mapsto V$ 
  with  $P(i + \lambda) = (1 - \lambda)P(i) + \lambda P(i+1)$
  for all $\lambda \in [0,1]$ and all $i \in \{0, 1, \dots, n-1\}$.
\end{mydef}

In the following we will focus on polygonal curves (polygonal chains) 
instead of classical continuous curves, ofen used in pure mathematics, 
because most data we have in this area is 
captured by some none continuous mechanism, e.g., GPS positions from a tracking device.
Addionally, we focus on the 2D Space, i.e. $V = \mathbb{R}^2$.
However, many concepts also apply to normal curves and higher dimensional spaces.

% Hausdorff distance
The commonly known \textit{Hausdorff distance} can be understood 
as the largest distance between any two points on the curves. 
For each curve, a point with the maximum shortest distance to a point on the other curve is searched for, see \autoref{fig_hausdorff_distance_example}. \cite{alt_computing_1995}

\begin{mydef}
  \label{def_hausdorff_distance}
  Given two (polygonal) curves $P$ and $Q$, 
  their Hausdorff distance $\delta_{hd}(P,Q)$ is mathematically defined as: 
  $$\delta_{hd}(P,Q) = \max \left( \sup_{p \in P} \inf_{q \in Q} \text{dist}(p,q), \sup_{q \in Q} \inf_{p \in P} \text{dist}(p,q) \right)$$
\end{mydef}

This metric is very intuitive and can be easily computed for polygonal curves \cite{ko_complexity_2013}.
It finds applications in computer vision and computer graphics 
where it is used to compare the similarity of two mesh objects \cite{cignoni_metro_1998,ko_complexity_2013}. 
However, this metric does not take the course of the curves into account, 
but only the point sets of the individual curves.
Two curves may have a small Hausdorff-distance but seem very dissimilar 
when looking at them, see \autoref{fig_hausdorff_distance_bad_example}.

@@

% Fréchet distance motivation
In reality, the course of the curves is important for many applications. 
For example, in handwriting character recognition on digital input devices, especially in Chinese, % TODO: add more citations
gesture recognition on smartphone touchscreens \cite{hu_research_2022}, % TODO: add more citations
and motion evaluation in real-time motion capturing \cite{qiao_real-time_2017, shehu_curve_2012}. % TODO: add more citations
For this reason, the \textit{Fréchet distance}, named after the famous french mathematician René Maurice Fréchet, was introduced.

A common analogy for the Fréchet distance can be shown by a man walking his dog. 
Assume that both the man and the dog are moving along a predefined fixed curve. 
Both are free to adjust their speed as they wish, even to stop. 
However, they are not allowed to go backwards. 
We are looking for the minimum length of the dog leash 
that is necessary for both of them to reach their destination. 
The length of the dog leash is then exactly the Fréchet distance of the two curves.

@@

\begin{mydef}
  \label{def_frechet_distance}
  Given two (polygonal) curves $P: [a,a'] \mapsto V$, $Q: [b,b'] \mapsto V$ and a distance function $\text{dist}: (V,V) \mapsto \mathbb{R}$, 
  their Fréchet distance $\delta_{F}(P,Q)$ is mathematically defined as: 
  $$\delta_{F}(P,Q) = \inf_{\substack{\alpha: [0,1] \mapsto [a, a'] \\ \beta: [0,1] \mapsto [b, b']}} \max_{t \in [0,1]} \text{dist}(P(\alpha(t)), Q(\beta(t)))$$
  where $\alpha$ and $\beta$ represend the walking speed as 
  continuous monotonically increasing functions with
  $\alpha(0) = a, \alpha(1) = a'$, 
  $\beta(0) =b, \beta(1) =b'$.
\end{mydef}

% TODO: Add example of Frechet distance

% Scope
% - Different Variants and their differences
% - Original Algorithm by Alt and Godau
% - Original Algorihm for discrete Fréchet distance
% - Newer Algorithms (vor variants)

% Fréchet distance Variants 
\section{Variants}
Through the years many slightly differnt variants of the continuous Fréchet distance (see Definition \ref{def_frechet_distance}) have evolved. 
The continuous Fréchet distance takes into account every single point on the curves, 
with interpolation in the case of polygonal curves, 
without allowing to walk backwards 
and without any obstacles in between the two curves.
